DATA 621 - Business Analytics and Data Mining

### 2020-12-03 Final_project/Final Project Paper - Group 2.Rmd
Update

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(jsonlite)
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(caret)
library(data.table)
```
```{r loadData}
currentPath <- getwd()
dataPath <- "Data2"
us_path <- file.path(currentPath, dataPath, "USvideos.csv")

@@

Given the models, we will compare the RMSE and MAE on a special holdout set and
will select the model that performs best as the winner for this paper. We do not
expect it to outperform more sophisticated models.

@@

@@

@@

A feature that we will also consider is `lkratio`: the ratio of likes to
dislikes. This will help ascertain if it is the magnitude or the preponderance
which has a greater effect. As there may be some rare videos with no dislikes,
the ratios will be set to 1. To capture the effect of likes with no dislikes, an
indicator variable will be created or those "perfect" videos.

```{r addFeatures}
usExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
                 lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
```

@@

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
seenSet <- usExtract[seenIDX, ]
hideSet <- usExtract[-seenIDX, ]
```

### Multiple Linear Regression
Linear regression may be the best known algorithm used when analyzing a
continuous numeric outcome. It searches for a linear relationship of the
predictors that minimizes the squared error between the "predictor" function and
the observations [@Sheather2009AMA]. This model will engage in stepwise feature
using the AIC as the optimization metric.

```{r modelLinear1}
# Using stepAIC means no cross-validation. Train on entire dataset.
trC <- trainControl(method = 'none')

# Capture the unique country codes
set.seed(181)
linMod1 <- train(views ~ likes + dislikes + comments + tags + category +
                   comments_disabled + ratings_disabled +
                   video_error_or_removed + perfect + lkratio,
                 data = usExtract, family = gaussian(link = 'identity'),
                 method = 'glmStepAIC', direction = 'both', trace = 0,
                 trControl = trC)
```

\footnotesize
```{r m1Table}
kable(summary(linMod1$finalModel)$coefficients,
      caption = "Model 1 Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

The sign of these coefficients make sense in the main. The most predictive
feature for views is `likes` followed by `dislikes`. Of course, one usually
views a video at least once prior to rating it. We set the baseline category to
`Entertainment`, considering it was the most popular. Therfore we expected
negative coefficients for other categories. Interestingly, `Film & Animation`
is positively correlated with likes over and above `Entertainment`, despite
having fewer views.

Other observations are that the magnitude of the coefficients is similar to that
of the intercept, indicating that the predictors provide large swings to final
views. The "worst" predictor by magnitude is the `Nonprofits & Activism`
category; its negative coefficient is more than twice the baseline intercept
views. The best, interestingly is if ratings are disabled. This is
counterintuitive at first, but is logical. The amount of `likes`, `dislikes`,
and `comments` are many orders of magnitude greater than 1. Therefore, their
coefficients can be much smaller. A Boolean variable is either 1 or 0, therefore
its coefficient is much greater even if its actual contribution is lower.

Variable importance can be seen from the table below. The further down the list
a coefficient preceded by a **`-`** is, the more important it is and the more it
affected the AIC.
\footnotesize
```
                                  Df   Deviance    AIC
<none>                               7.1243e+16 212209
+ `categoryAutos & Vehicles`       1 7.1222e+16 212209
+ tags                             1 7.1223e+16 212209
- `categoryPeople & Blogs`         1 7.1278e+16 212210
- categoryEducation                1 7.1278e+16 212210
- perfectTRUE                      1 7.1279e+16 212210
+ video_error_or_removedTRUE       1 7.1237e+16 212210
+ `categoryPets & Animals`         1 7.1239e+16 212210
+ categoryGaming                   1 7.1239e+16 212210
+ `categoryTravel & Events`        1 7.1240e+16 212210
+ comments_disabledTRUE            1 7.1240e+16 212210
+ categorySports                   1 7.1240e+16 212210
+ `categoryScience & Technology`   1 7.1241e+16 212211
+ categoryShows                    1 7.1243e+16 212211
- `categoryHowto & Style`          1 7.1287e+16 212211
- `categoryFilm & Animation`       1 7.1299e+16 212212
- `categoryNews & Politics`        1 7.1321e+16 212214
- categoryMusic                    1 7.1338e+16 212215
- `categoryNonprofits & Activism`  1 7.1345e+16 212216
- ratings_disabledTRUE             1 7.1463e+16 212227
- categoryComedy                   1 7.1467e+16 212227
- lkratio                          1 7.3555e+16 212413
- dislikes                         1 8.9091e+16 213650
- comments                         1 8.9807e+16 213701
- likes                            1 2.0217e+17 218939
```
\normalsize

@@

@@

<!-- Model Building -->
```{r holdOut}
```
```{r modelLinear1}
```
```{r m1Table}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(jsonlite)
library(knitr)
library(stringr)
library(ggplot2)
library(scales)
library(caret)
library(data.table)
```
```{r loadData}
currentPath <- getwd()
dataPath <- "Data2"
us_path <- file.path(currentPath, dataPath, "USvideos.csv")

@@

Given the models, we will compare the RMSE, \(R^2\), and MAE on a holdout set
and will select the model that performs best as the winner for this paper. We do
not expect it to outperform more sophisticated models.

@@

### Target Variable

```{r targDens}
ggplot(usSetN[measure == 'views'], aes(x = value)) +
  geom_density(kernel = "epanechnikov") +
  ggtitle("Target Density (semi-log scale)") +
  scale_x_log10(label = scientific)
```

On a log scale, `views` looks rather Gaussian, which implies it has a lognormal
distribution.

@@

As expected, the number of tags is orders of magnitude less than the rating or
comment variables. On average there are more likes than comments and more
comments then dislikes, but all three exhibit symmetric Gaussian-like behavior.

@@

We will consider the relationship between `views` and the numerical predictors
of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
`category`, whether or not comments or ratings were disabled and, if there was
an error with the video.

A feature that we will also consider is `lkratio`: the ratio of likes to
dislikes. This will help ascertain if it is the magnitude or the preponderance
which has a greater effect. As there may be some rare videos with no dislikes,
the ratios will be set to 1.

```{r addFeatures}
usExtract[, lkratio := ifelse(dislikes == 0, 1, likes / dislikes)]

modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
                   comments_disabled + ratings_disabled +
                   video_error_or_removed + lkratio,
                   data = usExtract, fullRank = TRUE)
modMat <- predict(modDum, usExtract)
```

@@

```{r holdOut}
set.seed(617)
seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
seenX <- modMat[seenIDX, ]
seenY <- usExtract$views[seenIDX]
hideX <- as.data.frame(modMat[-seenIDX, ])
hideY <- usExtract$views[-seenIDX]
```

### Simple Linear Regression
Linear regression may be the best known algorithm used when analyzing a
continuous numeric outcome. It searches for a linear relationship of the
predictors that minimizes the squared error between the "predictor" function and
the observations [@Sheather2009AMA]. This model will engage in stepwise feature
selection using the AIC as the optimization metric.

```{r lm1Train}
# Using stepAIC means no cross-validation. Train on entire dataset.
trC <- trainControl(method = 'none')
set.seed(181)
lm1 <- train(x = seenX, y = seenY, family = gaussian(link = 'identity'),
             method = 'glmStepAIC', direction = 'both', trace = 0,
             trControl = trC)
```

\footnotesize
```{r lm1Table}
kable(summary(lm1$finalModel)$coefficients,
      caption = "Model 1 Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

The signs of these coefficients make sense in the main. Increased `likes`,
`dislikes` are correlated with increased views. Of course one usually views a
video at least once prior to rating it. Increased `comments` is negatively
correlated with views.

We set the baseline category to `Entertainment`, considering it was the most
popular. Therefore we expected negative coefficients for other categories. We
were surprised that was not the case. The "worst" category predictor by
magnitude is the `Nonprofits & Activism`. The "best" is `Gaming`.

The factor predictors tend to have much higher magnitude coefficients than do
the numeric ones. This makes sense. The amount of `likes`, `dislikes`, and
`comments` are many orders of magnitude greater than 1. Therefore, their
coefficients can be much smaller. A Boolean variable is either 1 or 0, therefore
its coefficient is much greater even if its actual contribution is lower.

### Generalized Linear Model: Gaussian Errors & Log Link
The generalized linear model (GLM) is an extension of the simple linear model,
but the errors can be distributed per any member of the exponential family and
the relationship between some function of the predictors---called the link---and
the mean needs to be linear, not that the mean itself must be linear in the
predictors [@Faraway2006EtL].

The model under consideration here assumes a Gaussian distribution of the
errors, but a multiplicative relationship between the mean and the predictors.
This is expressed by using a log link function. This is **not** the canonical
link function for the Gaussian GLM, but as we are using numerical methods there
is no issue.

This approach *appears* similar to that of the common technique of performing a
standard linear regression on the logs of the observations, but is different. As
per [@Gelman2006Lta], one approach *"…log transforms observed values, while the*
*second one log transforms the expected value.…the key difference being the*
*relation between the predicted value and the variance."*

```{r lm2Train}
set.seed(181)
lm2 <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
             method = 'glmStepAIC', direction = 'both', trace = 0,
             trControl = trC)
```

\footnotesize
```{r lm2Table}
kable(summary(lm2$finalModel)$coefficients,
      caption = "Model 3 Linear Regression Output",
      digts = 3L, format.args = list(big.mark = ','))
```
\normalsize

### ElasticNet: Penalized Regression
Instead of using AIC to select features, one can make use of penalized
regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
which can perform feature selection. Using a squared error term, \(L_2\), is at
the heart of ridge regression [@HTF2009TEo]. Using both methods together is
called the elastic net [@ZH2005RaV]. To tune the hyperparameters, which includes
the weighting between Lasso and Ridge, we will use 10-fold cross-validation.

```{r lm3Train}
trC <- trainControl(method = 'cv', number = 10L)
tG <- expand.grid(fraction = seq(0.8, 1, 0.001),
                  lambda = seq(0, 0.02, 0.001))
set.seed(181)
lm3 <- train(x = seenX, y = seenY,
             method = 'enet', trControl = trC, tuneGrid = tG)
```

There is no clean table of coefficients with elasticNet. Rather there is a
sequence of models built behind the scenes. The coefficients of the selected
model can be found in the appendix.

@@

```{r modTest}
lm1P <- predict(lm1, newdata = hideX)
lm2P <- predict(lm2, newdata = hideX)
lm3P <- predict(lm3, newdata = hideX)
compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
                                 RMSE(lm3P, hideY)),
                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                               R2(lm2P, hideY, formula = 'traditional'),
                               R2(lm3P, hideY, formula = 'traditional')),
                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
                                MAE(lm3P, hideY)))
kable(compTable, digits = 3L, format.args = list(big.mark = ','),
      caption = "Model Performance on Test Set")
```

@@

<!-- Model Building -->
```{r holdOut}
```
```{r lm1Train}
```
```{r lm1Table}
```
```{r lm2Train}
```
```{r lm2Table}
```
```{r lm3Train}
```

<!-- Model Evaluation -->
```{r modTest}
```
